---
title: "Statistics and analyses"
output: html_notebook
---

```{r}
library(ggplot2)
library(grid)
library(gridExtra)
library(readr)
library(curl)
library(stringr)
library(RColorBrewer)
library(plyr)
```

# Count segments, worldwide and areally

##Prepare the data:

```{r}
segbo_data <- read_csv("bsd_with_glottolog-with-features.csv", col_types = cols(BorrowedSound = col_character()))
seg_table <- table(segbo_data$BorrowedSound)
seg_table  <- seg_table[order(seg_table, decreasing=F)]
df <- as.data.frame(seg_table)
languoids <- read_csv("languoid.csv")

# Download Glottolog 4.0 (22.10.2019) geo data
glottolog_data <- read_csv("https://cdstar.shh.mpg.de/bitstreams/EAEA0-5787-1678-EBFE-0/languages_and_dialects_geo.csv")

# Download PHOIBLE 2.0 data
temp <- tempfile()
download.file("https://github.com/phoible/dev/archive/v2.0.zip",temp)
phoible_data <- read_csv(unz(temp, "dev-2.0/data/phoible.csv"), col_types = cols(SpecificDialect = col_character()))
unlink(temp)
```

##Make worlwide plots:
```{r}
seg_lang_sum <- nrow(unique(subset(segbo_data,select=c(InventoryID)))) # Count number of SegBo doculects

cairo_pdf(filename = paste("world_segments.pdf",sep=""),  height=23/148*nrow(df), onefile = FALSE)
p <- ggplot(df, aes(x=Var1, y=Freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, max(df$Freq)*1.1) +
  coord_flip() + 
  # ggtitle('Worldwide') +
  xlab("Segment") + ylab("Counts")+
  geom_text(aes(label=paste(Freq," (",round(Freq/seg_lang_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

worldwide_top <- subset(df, Freq > 2)
cairo_pdf(filename = paste("world_segments_top.pdf",sep=""),  height=23/148*nrow(worldwide_top), onefile = FALSE)
p <- ggplot(worldwide_top, aes(x=Var1, y=Freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) + 
  ylim(0, 11*(max(worldwide_top$Freq)/10)) +
  coord_flip() + 
  # ggtitle('Worldwide') +
  xlab("Segment") + ylab("Counts")+
  geom_text(aes(label=paste(Freq," (",round(Freq/seg_lang_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p
```


##Make areal plots:
```{r}
for (area in (unique(segbo_data$macroarea))){
  if (!is.na(area)) {
    area_seg <- subset(segbo_data, macroarea == area, select=c(BorrowedSound))
    area_seg_table <- table(area_seg$BorrowedSound)
    area_seg_table  <- area_seg_table[order(area_seg_table, decreasing=F)]
    df <- as.data.frame(area_seg_table)
    area_seg_lang_sum <- nrow(unique(subset(segbo_data, macroarea == area,select=c(InventoryID)))) # Count number of SegBo doculects in the macroarea
    
    
    cairo_pdf(filename = paste("areal_segments_",gsub(" ", "_", area), ".pdf",sep=""),  height=23/148*nrow(df), onefile = FALSE)
    df2 <- aggregate(Freq ~ Var1, df, sum)
    p <- ggplot(df2, aes(x=Var1, y=Freq)) +
      geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
      ylim(0, 11*(max(df2$Freq)/10)) +
      coord_flip() +
      # ggtitle(area) +
      xlab("Segment") + ylab("Counts")+
      geom_text(aes(label=paste(Freq," (",round(Freq/area_seg_lang_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
    print(p)
    
    dev.off()
    p
  }
}
area = "Eurasia"
area_seg <- subset(segbo_data, macroarea == area, select=c(BorrowedSound))
area_seg_table <- table(area_seg$BorrowedSound)
area_seg_table  <- area_seg_table[order(area_seg_table, decreasing=F)]
df <- as.data.frame(area_seg_table)
area_seg_lang_sum <- nrow(unique(subset(segbo_data, macroarea == area,select=c(InventoryID)))) # Count number of SegBo doculects in the macroarea

eurasia_top1 <- subset(df, Freq > 1)
eurasia_top2 <- subset(df, Freq == 1)
eurasia_once_string <- toString(eurasia_top2$Var1)


cairo_pdf(filename = paste("areal_segments_",area, "1.pdf",sep=""),  height=23/148*nrow(eurasia_top1), onefile = FALSE)
p <- ggplot(eurasia_top1, aes(x=Var1, y=Freq)) +
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, 11*(max(eurasia_top1$Freq)/10)) +
  coord_flip() +
  # ggtitle(area) +
  xlab("Segment") + ylab("Counts")+
  geom_text(aes(label=paste(Freq," (",round(Freq/area_seg_lang_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

cairo_pdf(filename = paste("areal_segments_",area, "2.pdf",sep=""),  height=23/148*nrow(eurasia_top2), onefile = FALSE)
p <- ggplot(eurasia_top2, aes(x=Var1, y=Freq)) +
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, 11*(max(eurasia_top2$Freq)/10)) +
  coord_flip() +
  # ggtitle(area) +
  xlab("Segment") + ylab("Counts")+
  geom_text(aes(label=paste(Freq," (",round(Freq/area_seg_lang_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)+ ylim(0, 60)
print(p)
dev.off()
p

area = "Africa"
area_seg <- subset(segbo_data, macroarea == area, select=c(BorrowedSound))
area_seg_table <- table(area_seg$BorrowedSound)
area_seg_table  <- area_seg_table[order(area_seg_table, decreasing=F)]
df <- as.data.frame(area_seg_table)
area_seg_lang_sum <- nrow(unique(subset(segbo_data, macroarea == area,select=c(InventoryID)))) # Count number of SegBo doculects in the macroarea

africa_top1 <- subset(df, Freq > 1)
africa_top2 <- subset(df, Freq == 1)
africa_once_string <- toString(africa_top2$Var1)

cairo_pdf(filename = paste("areal_segments_",area, "1.pdf",sep=""),  height=23/148*nrow(africa_top1), onefile = FALSE)
p <- ggplot(africa_top1, aes(x=Var1, y=Freq)) +
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, 11*(max(africa_top1$Freq)/10)) +
  coord_flip() +
  # ggtitle(area) +
  xlab("Segment") + ylab("Counts")+
  geom_text(aes(label=paste(Freq," (",round(Freq/area_seg_lang_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

cairo_pdf(filename = paste("areal_segments_",area, "2.pdf",sep=""),  height=23/148*nrow(africa_top2), onefile = FALSE)
p <- ggplot(africa_top2, aes(x=Var1, y=Freq)) +
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, 11*(max(africa_top2$Freq)/10)) +
  coord_flip() +
  # ggtitle(area) +
  xlab("Segment") + ylab("Counts")+
  geom_text(aes(label=paste(Freq," (",round(Freq/area_seg_lang_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)+ ylim(0, 15)
print(p)
dev.off()
p

```


# Plot histogram of number of borrowed segments in languages
```{r}
lang_table <- table(segbo_data$InventoryID)
lang_table  <- lang_table[order(lang_table, decreasing=F)]
df <- as.data.frame(lang_table)
df <- merge(x = df, y = unique(segbo_data[,c("InventoryID","name")]), by.x = "Var1",by.y = "InventoryID", all.x = TRUE)
df <- df[order(df$Freq, decreasing=T),]
borrowing_sum <- nrow(segbo_data) # Count number of SegBo rows

df_hist <- as.data.frame(table(df$Freq))
cairo_pdf(filename = paste("world_lang_hist.pdf",sep=""),  height=40/148*nrow(df_hist), onefile = FALSE)
p <- ggplot(df_hist, aes(x=Var1, y=Freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  xlab("Number of borrowed segments") + ylab("Number of languages")+
  geom_text(aes(label=Freq), vjust=-0.5, size=3)
print(p)
dev.off()
p
```

# Make borrower plots (and print average and median number of borrowed segments in each macroarea):
```{r}
cairo_pdf(filename = paste("world_langs.pdf",sep=""),  height=23/148*nrow(df), onefile = FALSE)
# p <- ggplot(df, aes(x=reorder(Var1,Freq), y=Freq)) + 
p <- ggplot(df, aes(x=Var1, y=Freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  coord_flip() + 
  # ggtitle('Worldwide') +
  xlab("") +
  ylab("Number of borrowed segments")+
  geom_text(aes(label=Freq), vjust=0.5, hjust=-0.1, size=3) +
  scale_x_discrete(labels=rev(df$name))
# ylim(0, max(df$Freq)*1.1) +
# geom_text(aes(label=paste(Freq," (",round(Freq/borrowing_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

df2 <- subset(df, Freq > 6)

cairo_pdf(filename = paste("world_langs_half.pdf",sep=""),  height=23/148*nrow(df2), onefile = FALSE)
p <- ggplot(df2, aes(x=Var1, y=Freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  coord_flip() + 
  # ggtitle('Worldwide') +
  xlab("") +
  ylab("Number of borrowed segments")+
  geom_text(aes(label=Freq), vjust=0.5, hjust=-0.5, size=3)+
  scale_x_discrete(labels=rev(df2$name))
# ylim(0, max(df$Freq)*1.1) +
# geom_text(aes(label=paste(Freq," (",round(Freq/borrowing_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

area_seg_mean <- list()

for (area in (unique(segbo_data$macroarea))){
  if (!is.na(area)) {
    area_lang <- subset(segbo_data, macroarea == area, select=c(name,InventoryID))
    area_lang_table <- table(area_lang$InventoryID)
    area_lang_table  <- area_lang_table[order(area_lang_table, decreasing=F)]
    df <- as.data.frame(area_lang_table)
    df <- merge(x = df, y = unique(segbo_data[,c("InventoryID","name")]), by.x = "Var1",by.y = "InventoryID", all.x = TRUE)
    df <- df[order(df$Freq, decreasing=T),]
    cairo_pdf(filename = paste("areal_langs_",gsub(" ", "_", area), ".pdf",sep=""),  height=23/148*nrow(df), onefile = FALSE)
    p <- ggplot(df, aes(x=Var1, y=Freq)) + 
      geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
      coord_flip() + 
      # ggtitle(area) +
      xlab("") +
      ylab("Number of borrowed segments")+
      geom_text(aes(label=Freq), vjust=0.5, hjust=-0.5, size=3)+
      scale_x_discrete(labels=rev(df$name))
    print(p)
    dev.off()
    p
    
     # average number of borrowed segments in macroarea
    print(paste(area,"-",round(mean(area_lang_table),2)," - ",round(median(area_lang_table),2)))
    
    Freq_limit <- list(Eurasia = 5, Papunesia = 5, Africa = 4, "North America" = 4, "South America" = 4 , "Australia" = 3)
    df2 <- subset(df, Freq > Freq_limit[[area]])
    df2 <- df2[order(df2$Freq, decreasing=T),]
    
    cairo_pdf(filename = paste("areal_langs_",gsub(" ", "_", area), "_half.pdf",sep=""),  height=1+0.153*nrow(df2), onefile = FALSE)
    p <- ggplot(df2, aes(x=Var1, y=Freq)) + 
      geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
      coord_flip() + 
      # ggtitle(area) +
      xlab("") +
      ylab("Number of borrowed segments")+
      geom_text(aes(label=Freq), vjust=0.5, hjust=-0.1, size=3)+
      scale_x_discrete(labels=rev(df2$name))
    print(p)
    dev.off()
    p
    }
}

```


# Make worldwide donor plots:
```{r}
source_lang_cor <- cbind(segbo_data)
source_lang_cor$SourceLanguageGlottocode <- gsub('\\s+', '', source_lang_cor$SourceLanguageGlottocode)
source_lang_cor$SourceLanguageGlottocode <- gsub('arab1395', 'stan1318', source_lang_cor$SourceLanguageGlottocode)
source_lang_cor$SourceLanguageGlottocode <- gsub('macr1272', 'port1283', source_lang_cor$SourceLanguageGlottocode)
source_lang_cor$SourceLanguageGlottocode <- gsub('nort3191', 'alge1239', source_lang_cor$SourceLanguageGlottocode)
source_lang_cor$SourceLanguageGlottocode[is.na(source_lang_cor$SourceLanguageGlottocode)] <- "unknown"

source_lang_glts <- unique(unlist(strsplit(as.character(source_lang_cor$SourceLanguageGlottocode), ",")))
# source_lang_glts <- source_lang_glts[!is.na(source_lang_glts) & source_lang_glts != "unknown"]

source_langs <- data.frame(glottocode=as.character(cbind(source_lang_glts)), stringsAsFactors=FALSE)
source_langs$count <- rep(0,nrow(source_langs))
for (row in 1:nrow(source_langs)) {
  source_langs$count[row] <- sum(str_count(source_lang_cor$SourceLanguageGlottocode,source_langs$glottocode[row]), na.rm=T)
}

source_langs <- merge(source_langs,languoids[,c("name","id")],by.x = "glottocode",by.y = "id",all.x=TRUE)
source_langs$name[source_langs$glottocode=="unknown"] <- "Unknown"
source_langs  <- source_langs[order(source_langs$count, decreasing=T),]

borrowing_sum <- nrow(segbo_data) # Count number of SegBo rows

cairo_pdf(filename = paste("world_source_langs.pdf",sep=""),  height=max(2,23/148*nrow(source_langs)), onefile = FALSE)
p <- ggplot(source_langs, aes(x=reorder(name,count), y=count)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  coord_flip() + 
  # ggtitle('Worldwide') +
  xlab("Source language") + ylab("Number of borrowed segments")+
  # geom_text(aes(label=count), vjust=0.5, hjust=-0.1, size=3)
  ylim(0, max(source_langs$count)*1.15) +
  geom_text(aes(label=paste(count," (",round(count/borrowing_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p


df <- subset(source_langs,count>9)
cairo_pdf(filename = paste("world_source_langs_half.pdf",sep=""),  height=23/148*nrow(df), onefile = FALSE)
p <- ggplot(df, aes(x=reorder(name,count), y=count)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  coord_flip() + 
  # ggtitle('Worldwide') +
  xlab("Source language") + ylab("Number of borrowed segments")+
  # geom_text(aes(label=count), vjust=0.5, hjust=-0.1, size=3)
  ylim(0, max(df$count)*1.15) +
  geom_text(aes(label=paste(count," (",round(count/borrowing_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

sum(source_langs[2:6,]$count) # big-five impact
sum(source_langs[7:nrow(source_langs),]$count) # other languages' impact

```

# Areal donor plots:
```{r}
for (area in (unique(segbo_data$macroarea))){
  if (!is.na(area)) {
    source_lang_cor_area <- subset(source_lang_cor, macroarea == area)
    source_lang_glts <- unique(unlist(strsplit(as.character(source_lang_cor_area$SourceLanguageGlottocode), ",")))
    # source_lang_glts <- source_lang_glts[!is.na(source_lang_glts) & source_lang_glts != "unknown"]
    
    source_langs <- data.frame(glottocode=as.character(cbind(source_lang_glts)), stringsAsFactors=FALSE)
    source_langs$count <- rep(0,nrow(source_langs))
    for (row in 1:nrow(source_langs)) {
      source_langs$count[row] <- sum(str_count(source_lang_cor_area$SourceLanguageGlottocode,source_langs$glottocode[row]), na.rm=T)
    }
    
    source_langs <- merge(source_langs,languoids[,c("name","id")],by.x = "glottocode",by.y = "id",all.x=TRUE)
    source_langs$name[source_langs$glottocode=="unknown"] <- "Unknown"
    source_langs  <- source_langs[order(source_langs$count, decreasing=T),]
    
    borrowing_sum_area <- nrow(source_lang_cor_area) # Count number of SegBo rows
    
    cairo_pdf(filename = paste("areal_source_langs_",gsub(" ", "_", area),".pdf",sep=""),  height=max(23/148*nrow(source_langs),2), onefile = FALSE)
    p <- ggplot(source_langs, aes(x=reorder(name,count), y=count)) + 
      geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
      coord_flip() + 
      # ggtitle(area) +
      xlab("Source language") + ylab("Number of donated segments")+
      # geom_text(aes(label=count), vjust=0.5, hjust=-0.1, size=3)
      ylim(0, max(source_langs$count)*1.13) +
      geom_text(aes(label=paste(count," (",round(count/borrowing_sum_area*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
    print(p)
    dev.off()
    p
    
  }
}

```

# Plot of result of borrowing events:
```{r}
result_table <- table(segbo_data$Result)
result_table  <- result_table[order(result_table, decreasing=F)]
df <- as.data.frame(result_table)
seg_sum <- nrow(segbo_data) # Count number of SegBo doculects

cairo_pdf(filename = paste("world_result.pdf",sep=""),  height=2, onefile = FALSE)
p <- ggplot(df, aes(x=Var1, y=Freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, max(df$Freq)*1.12) +
  coord_flip() + 
  # ggtitle('Worldwide') +
  xlab("Result") + ylab("Counts")+
  geom_text(aes(label=paste(Freq," (",round(Freq/seg_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

for (area in (unique(segbo_data$macroarea))){
  if (!is.na(area)) {
    area_result <- subset(segbo_data, macroarea == area, select=c(Result))
    result_table <- table(area_result$Result)
    result_table  <- result_table[order(result_table, decreasing=F)]
    df <- as.data.frame(result_table)
    seg_sum <- nrow(area_result) # Count number of SegBo doculects
    
    cairo_pdf(filename = paste("areal_result_",gsub(" ", "_", area),".pdf",sep=""),  height=max(23/148*nrow(source_langs),2), onefile = FALSE)
    p <- ggplot(df, aes(x=Var1, y=Freq)) + 
      geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
      ylim(0, max(df$Freq)*1.12) +
      coord_flip() + 
      # ggtitle(area) +
      xlab("Result") + ylab("Counts")+
      geom_text(aes(label=paste(Freq," (",round(Freq/seg_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
    print(p)
    dev.off()
    p
  }
}
```

# New distinction plots
```{r}
dist_table <- table(segbo_data$NewDistinction)
dist_table  <- dist_table[order(dist_table, decreasing=F)]
df <- as.data.frame(dist_table)
seg_sum <- nrow(segbo_data) # Count number of SegBo doculects

cairo_pdf(filename = paste("world_dist.pdf",sep=""),  height=max(23/148*nrow(dist_table),2), onefile = FALSE)
p <- ggplot(df, aes(x=Var1, y=Freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, max(df$Freq)*1.12) +
  coord_flip() + 
  # ggtitle('Worldwide') +
  xlab("New Distinction") + ylab("Counts")+
  geom_text(aes(label=paste(Freq," (",round(Freq/seg_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

for (area in (unique(segbo_data$macroarea))){
  if (!is.na(area)) {
    area_dist <- subset(segbo_data, macroarea == area, select=c(NewDistinction))
    dist_table <- table(area_dist$NewDistinction)
    dist_table  <- dist_table[order(dist_table, decreasing=F)]
    df <- as.data.frame(dist_table)
    seg_sum <- nrow(area_dist) # Count number of SegBo doculects
    
    cairo_pdf(filename = paste("areal_dist_",gsub(" ", "_", area),".pdf",sep=""),  height=max(23/148*nrow(dist_table),2), onefile = FALSE)
    p <- ggplot(df, aes(x=Var1, y=Freq)) + 
      geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
      ylim(0, max(df$Freq)*1.12) +
      coord_flip() + 
      # ggtitle(area) +
      xlab("New Distinction") + ylab("Counts")+
      geom_text(aes(label=paste(Freq," (",round(Freq/seg_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
    print(p)
    dev.off()
    p
  }
}
```

# SegBo coverage in macroareas, compared with Glottolog:
```{r}
seg_coverage_df <- unique(segbo_data[c("InventoryID","macroarea")])
seg_coverage_table <- table(seg_coverage_df$macroarea)
glt_coverage_table <- table(glottolog_data[glottolog_data$level=="language",]$macroarea)
# Produce coverage stats:
"SegBo coverage:"
100*seg_coverage_table/(sum(seg_coverage_table))
"Glottolog coverage:"
100*glt_coverage_table/(sum(glt_coverage_table))
```

# Check inter-areal borrowing:
```{r}
same_macro <- 0
diff_macro <- 0
check_list <- list()

for (i in 1:nrow(source_lang_cor)){
  s_langs <- unlist(strsplit(as.character(source_lang_cor[i,]$SourceLanguageGlottocode), ","))
  if (source_lang_cor[i,]$SourceLanguageGlottocode!="unknown"){
    for (s_lang in s_langs){
      if (!identical(character(0),glottolog_data[glottolog_data$glottocode==s_lang,]$macroarea)){
        if ((glottolog_data[glottolog_data$glottocode==s_lang,]$macroarea==source_lang_cor[i,]$macroarea) & (!is.na(source_lang_cor[i,]$macroarea)) ){
          same_macro <- same_macro + 1
        } else {
          diff_macro <- diff_macro + 1
        }
      } else {
        check_list <- c(check_list,s_lang)
      }
    }
  }
}

same_macro
diff_macro
```

# Calculate areal and worldwide frequencies of PHOIBLE segments
```{r}
ph_seglist <- unique(phoible_data$Phoneme)
ph_seglist <- ph_seglist[!is.na(ph_seglist)]

freq_ph_df <- data.frame("Segment" = ph_seglist, stringsAsFactors=FALSE)

row.names(freq_ph_df) <- freq_ph_df$Segment

freq_ph_df <- freq_ph_df[-1]

areas <- list("Africa", "Australia", "Eurasia", "North America", "Papunesia", "South America")

phoible_with_macroarea <- merge(x=phoible_data, y=glottolog_data, by.x="Glottocode", by.y="glottocode", all.x=TRUE)

for (area in areas) {
  freq_ph_df[area] <- c(integer(nrow(freq_ph_df)))
  area_sub <- subset(phoible_with_macroarea, macroarea == area, select=c(Phoneme))
  counter <- as.data.frame(table(area_sub))
  lang_sum <- nrow(unique(subset(phoible_with_macroarea, macroarea == area, select=c(InventoryID))))
  counter$Freq <- counter$Freq/lang_sum
  row.names(counter) <- counter$area_sub
  # counter <- counter[ ,freq, drop=FALSE]
  for (seg in rownames(counter)){
    freq_ph_df[seg, area] <- counter[seg, 'Freq']
  }
}

freq_ph_df['Worldwide'] <- c(integer(nrow(freq_ph_df)))
area_sub <- subset(phoible_with_macroarea, select=c(Phoneme))
counter <- as.data.frame(table(area_sub))
counter <- counter[!is.na(counter$area_sub),]
lang_sum <- nrow(unique(subset(phoible_with_macroarea, select=c(InventoryID))))
counter$Freq <- counter$Freq/lang_sum
row.names(counter) <- counter$area_sub
# counter <- counter[ ,freq, drop=FALSE]
for (seg in rownames(counter)){
  freq_ph_df[seg, 'Worldwide'] <- counter[seg, 'Freq']
}
```


# Calculate areal and worldwide frequencies of SegBo segments
```{r}
segbo_seglist <- unique(segbo_data$BorrowedSound)

freq_segbo_df <- data.frame("Segment" = segbo_seglist, stringsAsFactors=FALSE)

row.names(freq_segbo_df) <- freq_segbo_df$Segment

freq_segbo_df <- freq_segbo_df[-1]

for (area in areas) {
  freq_segbo_df[area] <- c(integer(nrow(freq_segbo_df)))
  area_sub <- subset(segbo_data, macroarea == area, select=c(BorrowedSound))
  counter <- as.data.frame(table(area_sub))
  counter <- counter[!is.na(counter$area_sub),]
  lang_sum <- nrow(unique(subset(segbo_data, macroarea == area, select=c(InventoryID))))
  counter$Freq <- counter$Freq/lang_sum
  row.names(counter) <- counter$area_sub
  for (seg in rownames(counter)){
    freq_segbo_df[seg, area] <- counter[seg, 'Freq']
  }
}

freq_segbo_df['Worldwide'] <- c(integer(nrow(freq_segbo_df)))
area_sub <- subset(segbo_data, select=c(BorrowedSound))
counter <- as.data.frame(table(area_sub))
counter <- counter[!is.na(counter$area_sub),]
lang_sum <- nrow(unique(subset(segbo_data, select=c(InventoryID))))
counter$Freq <- counter$Freq/lang_sum
row.names(counter) <- counter$area_sub
for (seg in rownames(counter)){
  freq_segbo_df[seg, 'Worldwide'] <- counter[seg, 'Freq']
}
```

# Plot SegBo frequency vs. PHOIBLE frequency
```{r}
for (area in c('Worldwide',areas)) {
  cairo_pdf(paste("freq_scatter_",gsub(" ", "_", area), ".pdf",sep=""),height = 4)
  area_df <- data.frame("Segment" = segbo_seglist, stringsAsFactors=FALSE)
  row.names(area_df) <- area_df$Segment
  area_df <- area_df[-1]
  area_df['SegBo'] <- c(integer(nrow(area_df)))
  area_df[,'SegBo'] <- freq_segbo_df[,area]
  area_df[segbo_seglist,'PHOIBLE'] <- freq_ph_df[segbo_seglist,area]
  area_df[is.na(area_df[segbo_seglist,'PHOIBLE']),'PHOIBLE'] <- 0
  area_df <- subset(area_df, SegBo > 0)
  
  if (area == "Worldwide")  {
    x_title <- "Typological frequency"
    y_title <- "Frequency of borrowing"
  } else {
    x_title <- "Areal frequency"
    y_title <- "Areal frequency of borrowing"
  }
  
  p <- ggplot(area_df, aes(x=PHOIBLE, y=SegBo)) + #geom_point() + 
    # ggtitle(paste('Borrowing frequency vs. typological frequency:',area)) + 
    xlab(x_title) + ylab(y_title)+ 
    geom_text(label=rownames(area_df))+ xlim(0, 1)+ ylim(0, 0.5)
  
  print(p)
  dev.off()
  p
}
```


# Make plots with average lines
```{r}
for (area in c('Worldwide',areas)) {
  area_df <- data.frame("Segment" = segbo_seglist, stringsAsFactors=FALSE)
  row.names(area_df) <- area_df$Segment
  area_df <- area_df[-1]
  area_df['SegBo'] <- c(integer(nrow(area_df)))
  area_df[,'SegBo'] <- freq_segbo_df[,area]
  area_df[segbo_seglist,'PHOIBLE'] <- freq_ph_df[segbo_seglist,area]
  area_df[is.na(area_df[segbo_seglist,'PHOIBLE']),'PHOIBLE'] <- 0
  area_df <- subset(area_df, SegBo > 0)
  
  for (i in 1:nrow(area_df)) {
    if (area_df[i,"PHOIBLE"]>0.8) {
      area_df[i,"group"] <- "5"
    } else if (area_df[i,"PHOIBLE"]>0.6){
      area_df[i,"group"] <- "4"
    } else if (area_df[i,"PHOIBLE"]>0.4){
      area_df[i,"group"] <- "3"
    } else if (area_df[i,"PHOIBLE"]>0.2){
      area_df[i,"group"] <- "2"
    } else {
      area_df[i,"group"] <- "1"
    }
  }
  
  area_df.summary <- aggregate(. ~ group, mean, data=area_df)
  
  if (area == "Worldwide")  {
    x_title <- "Typological frequency"
    y_title <- "Frequency of borrowing"
  } else {
    x_title <- "Areal frequency"
    y_title <- "Areal frequency of borrowing"
  }
  
  
  cairo_pdf(paste("freq_scatter_w_lines_",gsub(" ", "_", area),".pdf",sep=""),height = 4)
  p <- ggplot(area_df, aes(x=PHOIBLE, y=SegBo, color = group)) + #geom_point() + 
    # ggtitle(paste('Borrowing frequency vs. typological frequency:',area)) + 
    xlab(x_title) + ylab(y_title)+ 
    geom_text(label=rownames(area_df))+ xlim(0, 1)+ ylim(0, 0.5) +
    geom_crossbar(data=area_df.summary, aes(ymin = SegBo, ymax = SegBo, x=(as.integer(group)*0.2)-0.1 ),
                  size=1,col=brewer.pal(n=8,name = "Dark2")[2:(length( unique(area_df$group) ) + 1)], width = .195) +
    theme(legend.position = "none") +
    scale_color_manual(values=brewer.pal(n=8,name = "Dark2")[2:(length( unique(area_df$group) ) + 1)])
  print(p)
  dev.off()
  p
}
```


# calculate borrowability factors
```{r}
bs_df <- data.frame("Segment" = segbo_seglist, stringsAsFactors=FALSE)
row.names(bs_df) <- bs_df$Segment
bs_df <- bs_df[-1]
bs_df['SegBo'] <- c(integer(nrow(bs_df)))
bs_df[,'SegBo'] <- freq_segbo_df[,'Worldwide']
bs_df[segbo_seglist,'PHOIBLE'] <- freq_ph_df[segbo_seglist,'Worldwide']
bs_df[is.na(bs_df[segbo_seglist,'PHOIBLE']),'PHOIBLE'] <- 0
bs_df <- subset(bs_df, SegBo > 0)

bs_df['b_S'] <- bs_df['SegBo']/(bs_df['PHOIBLE']-bs_df['PHOIBLE']^2)
bs_df <- subset(bs_df, PHOIBLE>0.005)
bs_df <- bs_df[order(bs_df$b_S,decreasing=F),]

bs_df1 <- bs_df

cairo_pdf(filename = paste("bs_plot_full",".pdf",sep=""),  height=max(23/148*nrow(bs_df1),2), onefile = FALSE)
p <- ggplot(bs_df1, aes(x=reorder(rownames(bs_df1),b_S), y=b_S)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, max(bs_df$b_S)*1.03) +
  coord_flip() + 
  xlab("Segment") + ylab("Borrowbility factor")+
  geom_text(aes(label=round(bs_df1$b_S,3)), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

lim1 <-  bs_df[2*round(nrow(bs_df)/3),"b_S"] # divide the long plot to three page-fitting ones
lim2 <-  bs_df[round(nrow(bs_df)/3),"b_S"]

bs_df1 <- subset(bs_df, b_S>=lim1)

cairo_pdf(filename = paste("bs_plot_most",".pdf",sep=""),  height=max(23/148*nrow(bs_df1),2), onefile = FALSE)
p <- ggplot(bs_df1, aes(x=reorder(rownames(bs_df1),b_S), y=b_S)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, max(bs_df$b_S)*1.03) +
  coord_flip() + 
  xlab("Segment") + ylab("Borrowbility factor")+
  geom_text(aes(label=round(bs_df1$b_S,3)), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p

bs_df1 <- subset(bs_df, (b_S<lim1) & (b_S>lim2))

cairo_pdf(filename = paste("bs_plot_mid",".pdf",sep=""),  height=max(23/148*nrow(bs_df1),2), onefile = FALSE)
p <- ggplot(bs_df1, aes(x=reorder(rownames(bs_df1),b_S), y=b_S)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, max(bs_df$b_S)*1.03) +
  coord_flip() + 
  xlab("Segment") + ylab("Borrowbility factor")+
  geom_text(aes(label=round(bs_df1$b_S,3)), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p
  
bs_df1 <- subset(bs_df, b_S<=lim2)

cairo_pdf(filename = paste("bs_plot_least",".pdf",sep=""),  height=max(23/148*nrow(bs_df1),2), onefile = FALSE)
p <- ggplot(bs_df1, aes(x=reorder(rownames(bs_df1),b_S), y=b_S)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, max(bs_df$b_S)*1.03) +
  coord_flip() + 
  xlab("Segment") + ylab("Borrowbility factor")+
  geom_text(aes(label=round(bs_df1$b_S,3)), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p
```

# Compare borrowability factors by feature:
```{r}
bs_df$segment <- rownames(bs_df)
features <- colnames(phoible_data)
features <- features[14:length(features)]

bs_w_features <- merge(x = bs_df, y = unique(phoible_data[,c("Phoneme",features)]), by.x = "segment", by.y = "Phoneme", all.x = TRUE)
rownames(bs_w_features) <- bs_w_features$segment

for (feat in features){
  for(value in unique(bs_w_features[,feat])){
    vec=bs_w_features[,feat]==value
    print(paste("Feature: ", feat,", value: ",value,", mean b_S: ",round(mean(bs_w_features[vec,]$b_S),3),sep=""))
  }
}

print(paste("vowels:",round(mean(subset(bs_w_features,((syllabic=="+")&(consonantal=="-")) )[,"b_S"]),3)))

print(paste("nasalized vowels:",round(mean(subset(bs_w_features,((syllabic=="+") & (consonantal=="-")) & (nasal=="+"))[,"b_S"]),3)))

print(paste("plain vowels:",round(mean(subset(bs_w_features,((syllabic=="+") & (consonantal=="-")) & (nasal=="-"))[,"b_S"]),3)))

print(paste("consonants:",round(mean(subset(bs_w_features,!((syllabic=="+")&(consonantal=="-")) )[,"b_S"]),3)))

print(paste("voiced consonants:",round(mean(subset(bs_w_features,!((syllabic=="+")&(consonantal=="-")) & (periodicGlottalSource=="+") )[,"b_S"]),3)))

print(paste("voiceless consonants:",round(mean(subset(bs_w_features,!((syllabic=="+")&(consonantal=="-")) & (periodicGlottalSource=="-") )[,"b_S"]),3)))

print(paste("liquids:",round(mean(subset(bs_w_features,((approximant=="+")&(consonantal=="+")) )[,"b_S"]),3)))

print(paste("nasals:",round(mean(subset(bs_w_features,((sonorant=="+")&(approximant=="-")) )[,"b_S"]),3)))

print(paste("fricatives:",round(mean(subset(bs_w_features,((continuant=="+")&(sonorant=="-")) )[,"b_S"]),3)))

print(paste("affricates:",round(mean(subset(bs_w_features,((delayedRelease=="+")&(continuant=="-")) )[,"b_S"]),3)))

print(paste("stops:",round(mean(subset(bs_w_features,((delayedRelease=="-")) )[,"b_S"]),3)))

print(paste("trills:",round(mean(subset(bs_w_features,((trill=="+")) )[,"b_S"]),3)))

print(paste("labiodentals:",round(mean(subset(bs_w_features,((labiodental=="+")) )[,"b_S"]),3)))

print(paste("laterals:",round(mean(subset(bs_w_features,((lateral=="+")) )[,"b_S"]),3)))

print(paste("dorsals:",round(mean(subset(bs_w_features,((dorsal=="+")) )[,"b_S"]),3)))

print(paste("coronals:",round(mean(subset(bs_w_features,((coronal=="+")) )[,"b_S"]),3)))
```

The following code might not run as an R script using RStudio on Windows.
One should run it using the console: Select all lines (Ctrl+A) and run selected (Ctrl+Enter)
This is a known issue in RStudio. See RStudio support:
"On Windows, R's source function does not work with files that include characters that aren't part of the current system encoding. You may have trouble with RStudio's Run All and Source on Save commands, as they rely on source." (https://support.rstudio.com/hc/en-us/articles/200532197-Character-Encoding)

# Rhotic study

## Define rhotics (r-like glyph, optinally with diacritics)
```{r}
rho_base <- "[rɾɹɽʀʁɻ]"
diacritics <- "[ʼ̥̬ʰ̹̜͈ ͓ ̟̠̩̯̈̽˞̤̰̼ʷʲˠˤ̴̝̞̘̙̪̺̻̃ⁿˡ̚ˀʱᶣ̊ ːˑ|]" # the character "|" in PHOIBLE indicates under-specified phonemes




```

## Find PHOIBLE segments that are rhotic
```{r}
is_rho_regexp <- paste("^(",diacritics, "*",rho_base, diacritics,"*",")+$",sep="")
ph_is_rho_bool <- grepl(is_rho_regexp, phoible_data$Phoneme)
ph_is_rho <- phoible_data[ph_is_rho_bool,]
phoible_data$is_rhotic <- ph_is_rho_bool # Add rhotic column to phoible data
```

## PHOIBLE rhotics
```{r}
ph_rho_counter <- count(phoible_data, c("InventoryID", "is_rhotic")) # Count rhotics and non-rhotics in doculects

ph_lang_sum <- nrow(unique(subset(phoible_data,select=c(InventoryID)))) # Count number of PHOIBLE doculects
ph_lang_rho_count <- count(ph_rho_counter$freq[ph_rho_counter$is_rhotic]) #  Histogram of doculects by number of rhotic segments they have (only languages with rhotics) 
ph_lang_rho_sum <- sum(ph_lang_rho_count$freq) # How many doculects have at least one rhotic
ph_hist_rho <- rbind(data.frame(x = 0, freq = ph_lang_sum-ph_lang_rho_sum), ph_lang_rho_count) # Add number of doculects without rhotics
ph_hist_rho$percent <- ph_hist_rho$freq/ph_lang_sum*100

cairo_pdf(filename = "rho_hist.pdf",  height=35/148*nrow(ph_hist_rho), onefile = FALSE)
p <- ggplot(ph_hist_rho, aes(x=reorder(x,-x), y=freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) + 
  ylim(0, 11.5*ceiling(max(ph_hist_rho$freq)/10)) +
  coord_flip() + 
  # ggtitle(paste(toString(nrow(ph_rhotics_to_plot)),'most frequent rhotic segments in PHOIBLE')) + 
  xlab("Number of rhotics") + ylab("Languages")+
  geom_text(aes(label=paste(freq," (",round(ph_hist_rho$percent,digits = 2),"%)",sep = "")), vjust=0.5, hjust=-0.17, size=3)
print(p)
dev.off()
p

ph_rhotics <- count(phoible_data$Phoneme[phoible_data$is_rhotic==TRUE]) # What rhotics are in phoible and how frequent they are
row.names(ph_rhotics) <- ph_rhotics$x
ph_rhotics <- ph_rhotics[order(ph_rhotics$freq),]

# If you want to see in different areas, you have to work with the phoible w/ macroarea

ph_rhotics_to_plot <- subset(ph_rhotics, freq>=5)
cairo_pdf(filename = "phoible_rhotics.pdf",  height=35/148*nrow(ph_rhotics_to_plot), onefile = FALSE)
p <- ggplot(ph_rhotics_to_plot, aes(x=reorder(x, freq), y=freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) + 
  ylim(0, 11.5*floor(max(ph_rhotics_to_plot$freq)/10)) +
  coord_flip() + 
  # ggtitle(paste(toString(nrow(ph_rhotics_to_plot)),'most frequent rhotic segments in PHOIBLE')) + 
  xlab("Segment") + ylab("Frequency")+
  geom_text(aes(label=paste(freq," (",round(freq/ph_lang_sum*100,digits = 1),"%)",sep = "")), vjust=0.5, hjust=-0.17, size=3)
print(p)
dev.off()
p

```

## SegBo rhotics
```{r}
seg_is_rho_bool <- grepl(is_rho_regexp, segbo_data$BorrowedSound) # Find SegBo segments that ARE rhotic (and nothing else, but diacritics)
seg_is_rho <- segbo_data[seg_is_rho_bool,]

segbo_data$is_rhotic <- seg_is_rho_bool # Add rhotic column to phoible data


seg_rho_counter <- count(segbo_data, c("InventoryID", "is_rhotic")) # Count rhotics and non-rhotics in doculects

seg_lang_sum <- nrow(unique(subset(segbo_data,select=c(InventoryID)))) # Count number of SegBo doculects
seg_lang_rho_count <- count(seg_rho_counter$freq[seg_rho_counter$is_rhotic]) #  Histogram of doculects by number of rhotic segments they have (only languages with rhotics) 
seg_lang_rho_sum <- sum(seg_lang_rho_count$freq) # How many doculects have at least one rhotic
seg_hist_rho <- rbind(data.frame(x = 0, freq = seg_lang_sum-seg_lang_rho_sum), seg_lang_rho_count) # Add number of doculects without rhotics
seg_hist_rho$percent <- seg_hist_rho$freq/seg_lang_sum*100

seg_rhotics <- count(segbo_data$BorrowedSound[segbo_data$is_rhotic==TRUE]) # What rhotics are in SegBo and how frequent they are
row.names(seg_rhotics) <- seg_rhotics$x
seg_rhotics <- seg_rhotics[order(-seg_rhotics$freq),]

cairo_pdf(filename = "segbo_rhotics.pdf",  height=35/148*nrow(seg_rhotics), onefile = FALSE)
p <- ggplot(seg_rhotics, aes(x=reorder(x, freq), y=freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) + 
  ylim(0, 12*floor(max(seg_rhotics$freq)/10)) +
  coord_flip() + 
  xlab("Segment") + ylab("Frequency")+
  geom_text(aes(label=paste(freq," (",round(freq/seg_lang_sum*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p
```

## Analysis of rhotic borrowing was done manually
```{r}
new_rho <- subset(segbo_data, is_rhotic & (Result != "other distributional change"))
rho_borrowers <- unique(new_rho$BorrowingLanguageGlottocode)
write_csv(new_rho,"new_rho.csv") # Export rhotic borrowings data

new_rho_analysis <- read_csv("new_rho_analysis.csv") # Import manually analysed data
```

## Rhotic statistics
```{r}
table(new_rho_analysis[new_rho_analysis$to_analysis=='Yes',]$macroarea)/table(segbo_data$macroarea)[row.names(table(new_rho_analysis$macroarea))]*100 # Frequency of rhotic brrowing within each macroarea
```

## Rhotic donors plot
```{r}
rho_source <- new_rho_analysis[new_rho_analysis$to_analysis=='Yes',]$SourceLanguageGlottocode
rho_source <- gsub('\\s+', '', rho_source)
rho_source <- gsub('arab1395', 'stan1318', rho_source)
rho_source <- gsub('macr1272', 'port1283', rho_source)
rho_source <- gsub('nort3191', 'alge1239', rho_source)
rho_source[is.na(rho_source)] <- "unknown"
rho_source_table <- table(unlist(strsplit(as.character(rho_source), ",")))
rho_source_table  <- rho_source_table[order(rho_source_table, decreasing=F)]

rho_source_glt <- row.names(rho_source_table)
rho_source_names <- list()
for (lang in rho_source_glt) {
  if (lang!="unknown") {
    rho_source_names <- append(rho_source_names,languoids[match(lang,languoids$id),]$name)
  }
  else {
    rho_source_names <- append(rho_source_names,"Unknown")
  }
}
rho_source_names <- as.character(rho_source_names)
rho_source_df <- data.frame("name" = rho_source_names, "Freq" = as.integer(rho_source_table))

cairo_pdf(filename = paste("rho_donors.pdf",sep=""),  height=23/148*nrow(rho_source_df), onefile = FALSE)
p <- ggplot(rho_source_df, aes(x=reorder(name,Freq), y=Freq)) + 
  geom_bar(position = "dodge" ,stat="identity", width = 0.5) +
  ylim(0, max(rho_source_df$Freq)*1.1) +
  coord_flip() + 
  # ggtitle("Rhotic donors") + 
  xlab("Donor language") + ylab("Donated Segments")+
  geom_text(aes(label=paste(Freq," (",round(Freq/sum(rho_source_table)*100,digits=1),"%)",sep="")), vjust=0.5, hjust=-0.1, size=3)
print(p)
dev.off()
p
```

